---
title: "Basic CVD model in DES simulation"
author: "Astrid Yu"
date: `r Sys.Date()`
output: html_document
---

```{r inputs}
source("life_death_model.Rmd")
inputs = list(
  ## basic/general from Stroke paper, 2023
  vN = 1000,
  vHorizon = 10,
  CPI2017 = 245.12,
  CPI2020 = 258.811,
  
  ## events from paper 2017
  mortalityAfterASCVD = 1.9,
  mortalityFirstYearASCVD_female_young = 0.09,
  mortalityFirstYearASCVD_female_old = 0.3,
  mortalityFirstYearASCVD_male_young = 0.14,
  mortalityFirstYearASCVD_male_old = 0.25,

  ## utility
  uHealthy = 1,
  uAfterASCVD = 0.773,
  uDeath = 0,
  
  ## cost(2017)
  cAnnualFU_afterASCVD= 3917,
  cNonFatalASCVD = 49348,
  cFatalASCVD = 16760,
  ### source：Agency for Healthcare Research and Quality. Mean expenditure per person by age groups, United States, 1996 to 2020. Medical Expenditure Panel Survey.(2019)
  cHealthcareNonCVD_18_44 = 3891,
  cHealthcareNonCVD_45_64 = 8494,
  cHealthcareNonCVD_65 = 13016

)
## inflation
inputs$rInflation = inputs$CPI2020 / inputs$CPI2017

## leave some spaces for PSA
```

# gomepertz model
```{r hazard-gome}

life_female_afterASCVD = data.frame(Year = rep(2010,111), Age = 0:110, qx = inputs$mortalityAfterASCVD * US_life_female_2010$qx, lx = rep(0,111), dx = rep(NA,111))
life_female_afterASCVD$lx[1] = 100000
for (i in 1:nrow(life_female_afterASCVD)) {
  if (life_female_afterASCVD[i, "lx"] > 0) {
    life_female_afterASCVD[i, "dx"] = round(life_female_afterASCVD[i, "qx"] * life_female_afterASCVD[i, "lx"])
    life_female_afterASCVD[i+1, "lx"] = life_female_afterASCVD[i, "lx"] - life_female_afterASCVD[i, "dx"]
  }
}
life_female_afterASCVD  = life_female_afterASCVD %>% filter(!is.na(dx))

dx_female_afterASCVD = data.frame(Age = rep(NA, 100000), Death = rep(1, 100000))

start = 1
for (age in 0:(nrow(life_female_afterASCVD)-1)) {
  n_row = life_female_afterASCVD[age+1,"dx"]
  dx_female_afterASCVD[seq(start,(n_row + start - 1)),"Age"] = age
  start = start + n_row
}

parameters_f = data.frame(
  Age = life_female_afterASCVD$Age,
  shape=rep(NA,nrow(life_female_afterASCVD)),
  rate=rep(NA,nrow(life_female_afterASCVD)))

for (age in 0:nrow(life_female_afterASCVD)-1) {
  surv.data <- with(dx_female_afterASCVD[dx_female_afterASCVD$Age >= age,], Surv(Age, Death, origin=age))
  surv.model <- flexsurvreg(surv.data ~ 1, dist="gompertz")
  
  parameters_f$shape[age+1] <- surv.model$coefficients[1]
  parameters_f$rate[age+1] <- exp(surv.model$coefficients[2])
}

#### male ####


life_male_afterASCVD = data.frame(Year = rep(2010,111), Age = 0:110, qx = inputs$mortalityAfterASCVD * US_life_male_2010$qx, lx = rep(0,111), dx = rep(NA,111))
life_male_afterASCVD$lx[1] = 100000
for (i in 1:nrow(life_male_afterASCVD)) {
  if (life_male_afterASCVD[i, "lx"] > 0) {
    life_male_afterASCVD[i, "dx"] = round(life_male_afterASCVD[i, "qx"] * life_male_afterASCVD[i, "lx"])
    life_male_afterASCVD[i+1, "lx"] = life_male_afterASCVD[i, "lx"] - life_male_afterASCVD[i, "dx"]
  }
}
life_male_afterASCVD  = life_male_afterASCVD %>% filter(!is.na(dx))

dx_male_afterASCVD = data.frame(Age = rep(NA, 100000), Death = rep(1, 100000))

start = 1
for (age in 0:(nrow(life_male_afterASCVD)-1)) {
  n_row = life_male_afterASCVD[age+1,"dx"]
  dx_male_afterASCVD[seq(start,(n_row + start - 1)),"Age"] = age
  start = start + n_row
}

parameters_m = data.frame(
  Age = life_male_afterASCVD$Age,
  shape=rep(NA,nrow(life_male_afterASCVD)),
  rate=rep(NA,nrow(life_male_afterASCVD)))

for (age in 0:nrow(life_male_afterASCVD)-1) {
  surv.data <- with(dx_male_afterASCVD[dx_male_afterASCVD$Age >= age,], Surv(Age, Death, origin=age))
  surv.model <- flexsurvreg(surv.data ~ 1, dist="gompertz")
  
  parameters_m$shape[age+1] <- surv.model$coefficients[1]
  parameters_m$rate[age+1] <- exp(surv.model$coefficients[2])
}

gompertz.afterASCVD.parameters = list(Age = 0:(nrow(parameters_f)-2), male.shape = parameters_m[1:(nrow(parameters_m)-1),"shape"], male.rate = parameters_m[1:(nrow(parameters_m)-1),"rate"], female.shape = parameters_f[1:(nrow(parameters_f)-1),"shape"], female.rate = parameters_f[1:(nrow(parameters_f)-1),"rate"])

ageAtDeath_afterASCVD <- function(currentAge, gender)
{
  currentAge_f <- floor(currentAge) #negative time to secular death issue
  
  # Male 1 FEMALE 2
  if(gender == 2)
  {
    shape <- gompertz.afterASCVD.parameters$female.shape[currentAge_f+1]
    rate  <- gompertz.afterASCVD.parameters$female.rate[currentAge_f+1]
  }
  else
  {
    shape <- gompertz.afterASCVD.parameters$male.shape[currentAge_f+1]
    rate  <- gompertz.afterASCVD.parameters$male.rate[currentAge_f+1]
  }
  
  min(currentAge + rgompertz(1, shape, rate), 105)
}
save(gompertz.parameters, gompertz.afterASCVD.parameters, file = "./life table/gompertz. RData")
```

# get CVD based on PCE

https://github.com/spgarbet/desdt/tree/main/framingham

```{r Pooled-Cohort}
source('pcr.R')
# # the probability are percents!!
# 
# # simply change from prob to rate than to event
ProbToRate = function(prob, t){
  -log(1-prob)/t
}
# # what this t should be -- 10 years?
# # time unit: year
# rate = ProbToRate(pcr()/100)
# days = rexp(1,rate)
```

# Sample population

```{r nhanes-clearance}
## SEQN: Respondent sequence number
# Examination: Body Measures
## BMXBMI - Body Mass Index (kg/m**2)
## 12.3 to 86.2	Range of Values
## .	Missing
nhanes_data_1 = data.frame(read_xpt("./NHANES/BMX_J.XPT", col_select = c("SEQN","BMXBMI")))
# Lab: Glycohemoglobin
## LBXGH: Glycohemoglobin (%)
nhanes_data_2 = data.frame(read_xpt("./NHANES/GHB_J.XPT", col_select = c("SEQN","LBXGH")))
# Examination: Blood Pressure
## BPXSY1 - Systolic: Blood pres (1st rdg) mm Hg
## BPXDI1 - Diastolic: Blood pres (1st rdg) mm Hg
## 0 to 136	Range of Values
## .	Missing
nhanes_data_3 = data.frame(read_xpt("./NHANES/BPX_J.XPT", col_select = c("SEQN","BPXDI1","BPXSY1")))
# Demographic Variables and Sample Weights
## RIAGENDR - Gender
## 1	Male
## 2	Female	
## .	Missing
## RIDAGEYR - Age in years at screening
## 0 to 79	Range of Values
## 80	80 years of age and over
## .	Missing

## RIDRETH1 - Race/Hispanic origin
## 1	Mexican American	1367	1367	
## 2	Other Hispanic	820	2187	
## 3	Non-Hispanic White	3150	5337	
## 4	Non-Hispanic Black	2115	7452	
## 5	Other Race - Including Multi-Racial	1802	9254	
## .	Missing

## WTMEC2YR - Full sample 2 year MEC exam weight
## 2566.1838545 to 419762.83649	Range of Values
## 0	Not MEC Examined
## .	Missing
nhanes_data_4 = data.frame(read_xpt("./NHANES/DEMO_J.XPT", col_select = c("SEQN","RIDAGEYR","RIAGENDR","RIDRETH1","WTMEC2YR"))) #WTMEC2YR is the sample weights variable
# Questionnaire: Diabetes
## DIQ010 - Doctor told you have diabetes
## 1	Yes
## 2	No
## 3	Borderline
## 7	Refused
## 9	Don't know
## .	Missing
nhanes_data_5 = data.frame(read_xpt("./NHANES/DIQ_J.XPT", col_select = c("SEQN","DIQ010")))
# Lab: Cholesterol - High - Density Lipoprotein
## LBDHDD: Direct HDL-Cholesterol (mg/dL)
nhanes_data_6 = data.frame(read_xpt("./NHANES/HDL_J.XPT", col_select = c("SEQN","LBDHDD")))
# Questionnaire: medical conditions
## MCQ160d: Ever told you had angina/angina pectoris
## MCQ160e: Ever told you had heart attack
## MCQ160f: Ever told you had a stroke
## 1	Yes
## 2	No
## 7	Refused	
## 9	Don't know
## .	Missing
nhanes_data_7 = data.frame(read_xpt("./NHANES/MCQ_J.XPT", col_select = c("SEQN","MCQ160D","MCQ160F","MCQ160E")))
# Questionnaire: Smoking - Cigarette Use
## SMQ040: Do you now smoke cigarettes?
## 1	Every day
## 2	Some days
## 3	Not at all	
## 7	Refused
## 9	Don't know
## .	Missing
nhanes_data_8 = data.frame(read_xpt("./NHANES/SMQ_J.XPT", col_select = c("SEQN","SMQ040")))
# Lab: Cholesterol - Total
## LBXTC: Total Cholesterol (mg/dL)
nhanes_data_9 = data.frame(read_xpt("./NHANES/TCHOL_J.XPT", col_select = c("SEQN","LBXTC")))
nhanes_data_10 = data.frame(read_xpt("./NHANES/BPQ_J.XPT", col_select = c("SEQN", "BPQ050A")))

nhanes_data_raw <- list(nhanes_data_1,nhanes_data_2,nhanes_data_3,nhanes_data_4,
                        nhanes_data_5,nhanes_data_6,nhanes_data_7,nhanes_data_8,nhanes_data_9,
                        nhanes_data_10)
nhanes_data_raw <- data.frame(reduce(nhanes_data_raw, full_join, by='SEQN'))
nhanes_pop = nhanes_data_raw %>%
  # whether needs to check missing data
  filter(WTMEC2YR != 0 & RIDAGEYR >= 40 & RIDAGEYR <= 79 & LBXTC >= 30 & LBXTC <= 500 & LBDHDD >= 5 & LBDHDD <= 200 & BPXSY1 >= 60 & BPXSY1 <= 200) %>%
  drop_na() %>%
  mutate(RIDRETH1 = ifelse(RIDRETH1 == 3, 1, ifelse(RIDRETH1 == 4, 2, ifelse(RIDRETH1 %in% c(1,2), 3, 4))))

race = 1:4
names(race) = c("White", "African American", "Hispanic", "Other")
race_convert = function(num) {
  return(names(race[num]))
}

bp_treatment = 1:2
names(bp_treatment) = c("Yes","No")

smoker = 1:3
names(smoker) = c("Every day","Some days","Not at all")

diabetic = 1:2
names(diabetic) = c("Yes","No")

save(nhanes_pop, nhanes_data_raw, file = "./NHANES/nhanes_full.RData")
```

```{r get-pop}
initialize_patient <- function(traj, inputs)
{
        traj %>%
                seize("time_in_model")       %>%
                # NHANES version
                set_attribute("aNo",         function() sample(1:nrow(nhanes_pop), 1, prob=1/nhanes_pop$WTMEC2YR)) %>%
                set_attribute("aGender",     function() nhanes_pop$RIAGENDR[get_attribute(env, "aNo")]) %>%
                set_attribute("aAge",        function() nhanes_pop$RIDAGEYR[get_attribute(env, "aNo")]) %>% 
                set_attribute("aAgeInitial", function() get_attribute(env, "aAge")) %>%
                set_attribute("aRace",       function() nhanes_pop$RIDRETH1[get_attribute(env, "aNo")]) %>%
                set_attribute("aTotChol",    function() nhanes_pop$LBXTC[get_attribute(env, "aNo")]) %>%
                set_attribute("aHdlChol",    function() nhanes_pop$LBDHDD[get_attribute(env, "aNo")]) %>%
                set_attribute("aSystolicBp", function() nhanes_pop$BPXSY1[get_attribute(env, "aNo")]) %>%
                set_attribute("aBpTreatment",function() nhanes_pop$BPQ050A[get_attribute(env, "aNo")]) %>%
                set_attribute("aSmoker",      function() nhanes_pop$SMQ040[get_attribute(env, "aNo")]) %>%
                set_attribute("aDiabetic",    function() nhanes_pop$DIQ010[get_attribute(env, "aNo")]) %>%
                set_attribute("aPrsZ",       function() 0)
}
```

# DES

```{r events-no-longer-used}
counters = c("get_CVD", "death_without_CVD", "death_with_CVD", "time_in_model")

terminate <- function(traj, inputs)
{
  traj %>%
  branch(function() 1,
         continue=FALSE,
         trajectory()               %>%
           release("time_in_model") %>%
           timeout(0)
        )
}

years_till_death_without_CVD <- function(inputs)
{
  age       <- get_attribute(env, 'aAge')
  gender    <- get_attribute(env, 'aGender')
  
  deathAge = ageAtDeath(age, gender)
  
  return(deathAge - age)
}

event_death_without_CVD <- function(traj, inputs)
{
  cat("event_Death() ", now(env), "\n")

  traj %>% 
    set_attribute("aAgeDeathNonCVD", function() get_attribute(env,"aDeathWithoutCVD") + get_attribute(env,"aAgeInitial")) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death without CVD") %>%
      mark("death_without_CVD") %>%
      terminate(inputs)
  )
}

years_till_get_CVD <- function(inputs)
{
  gender = ifelse(get_attribute(env, "aGender") == 1, "M", "F")
  race_pcr = race_convert(get_attribute(env, "aRace")) 
  
  prob = pcr(gender,
             get_attribute(env,"aAgeInitial") + now(env),
             race_pcr,
             get_attribute(env, "aTotChol"),
             get_attribute(env, "aHdlChol"),
             get_attribute(env, "aSystolicBp"),
             (get_attribute(env, "aBpTreatment") == 1),
             (get_attribute(env, "aSmoker") %in% c(1,2)),
             (get_attribute(env, "aDiabetic") == 1),
             get_attribute(env, "aPrsZ")
             )/100
  
  # 10-year probability to 1-year rate
  years = rexp(1,ProbToRate(prob, 10))
  
  return(years)
}

event_get_CVD = function(traj, inputs) {
  cat("event_get_CVD() ", now(env), "\n")

  traj %>%
    set_attribute("aAgeCVD", function() get_attribute(env,"aAgeInitial") + get_attribute(env, "aGetCVD")) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Get CVD") %>%
      mark("get_CVD")) %>% 
    timeout(function() years_till_death_with_CVD(inputs)) %>%
    event_death_with_CVD(inputs)
}

# years_till_acute_CVD_death = function(inputs) {
#   
#     
#   
#   return(ifelse(happen == 1, 1, 500))
#   }
# 
# event_death_acute_CVD = function(traj, inputs) {
#   traj %>% 
#     set_attribute("aAgeDeath", function() 1 + get_attribute(env,"aAgeCVD")) %>%
#     branch(
#     function() 1,
#     continue=c(TRUE),
#     trajectory("Death Acute CVD") %>%
#       mark("death Acute CVD") %>%
#       terminate(inputs)
#     )
#   }

# years_till_death_with_CVD <- function(inputs)
# {
#   ageCVD    <- get_attribute(env, 'aAgeCVD')
#   gender    <- get_attribute(env, 'aGender')
#   
#   # whether ASCVD died in first year happened
#   happen = 0
#   if (gender == 1) {
#     ## men
#     if (ageCVD < 65) {
#       happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_young, inputs$mortalityFirstYearASCVD_male_young))
#     } else {
#       happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_old, inputs$mortalityFirstYearASCVD_male_old))
#       }
#     } else {
#     ## women
#       if (ageCVD < 65) {
#         happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_young, inputs$mortalityFirstYearASCVD_female_young))
#       } else {
#         happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_old, inputs$mortalityFirstYearASCVD_female_old))
#       }
#     }
#   
#   deathAge = deathAge = ageAtDeath_afterASCVD(
#     ifelse(ageCVD+1>length(gompertz.afterASCVD.parameters$male.shape)-1, length(gompertz.afterASCVD.parameters$male.shape)-1, ageCVD), 
#   gender)
#   
#   return(ifelse(happen == 1, get_attribute(env, "aGetCVD") + 1, get_attribute(env, "aGetCVD") + deathAge - ageCVD))
# }

years_till_death_with_CVD <- function(inputs)
{
  ageCVD    <- get_attribute(env, 'aGetCVD') + get_attribute(env, "aAgeInitial")
  gender    <- get_attribute(env, 'aGender')
  
  ## getCVD maybe sth really big.
  
  age_max = ifelse(gender == 1, length(gompertz.afterASCVD.parameters$male.shape), length(gompertz.afterASCVD.parameters$female.shape))
  
  if (ageCVD >= age_max - 1) {
    # not able to use ageAtDeath(ageCVD+1, gender)
    years = get_attribute(env, 'aGetCVD')
  } else {
    ## ageCVD under the age limitation -- able to use gomepertz
    # whether ASCVD died in first year happened
    happen = 0
    if (gender == 1) {
      ## men
      if (ageCVD < 65) {
        happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_young, inputs$mortalityFirstYearASCVD_male_young))
        } else {
          happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_old, inputs$mortalityFirstYearASCVD_male_old))
          }
      } else {
        ## women
        if (ageCVD < 65) {
          happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_young, inputs$mortalityFirstYearASCVD_female_young))
          } else {
            happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_old, inputs$mortalityFirstYearASCVD_female_old))
          }
        }

  deathAge = ageAtDeath_afterASCVD(ageCVD + 1, gender)
  
  # GetCVD + ageInitial maybe a impossible number
  years = ifelse(happen == 1, get_attribute(env, "aGetCVD") + 1, deathAge - get_attribute(env, "aAgeInitial"))
  }
  
  return(years)
}

event_death_with_CVD = function(traj, inputs) {
  traj %>% 
    set_attribute("aAgeDeathCVD", function() get_attribute(env, "aAgeInitial")+now(env)) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death With CVD") %>%
      mark("death_with_CVD") %>%
      terminate(inputs)
    )
}

event_registry <- list(
        list(name          = "get CVD",
             attr          = "aGetCVD",
             time_to_event = years_till_get_CVD,
             func          = event_get_CVD,
             reactive      = FALSE),
        list(name          = "Death without CVD",
             attr          = "aDeathWithoutCVD",
             time_to_event = years_till_death_without_CVD,
             func          = event_death_without_CVD,
             reactive      = FALSE),
        # list(name          = "Death with CVD",
        #      attr          = "aDeathWithCVD",
        #      time_to_event = years_till_death_with_CVD,
        #      func          = event_death_with_CVD,
        #      reactive      = FALSE),
        list(name          = "Halt at time horizon",
             attr          = "aTerminate",
             time_to_event = function(inputs) inputs$vHorizon,
             func          = terminate,
             reactive      = FALSE)
)

```

```{r events2}
counters = c("get_CVD", "death_without_CVD", "death_of_ASCVD", "death_after_ASCVD", "time_in_model")

### Note:
### As only simulate all the things once, lots of places I use the initial age instead of the updated age. 

terminate <- function(traj, inputs)
{
  traj %>%
  branch(function() 1,
         continue=FALSE,
         trajectory()               %>%
           release("time_in_model") %>%
           timeout(0)
        )
}

years_till_death_without_CVD <- function(inputs)
{
  age       <- get_attribute(env, 'aAge')
  gender    <- get_attribute(env, 'aGender')
  
  deathAge = ageAtDeath(age, gender)
  
  ## the if-else here make sure that if get CVD firstly, will never die of 1.0 BG mortality
  
  return(ifelse(get_attribute(env, "aGetCVD") < deathAge - age, Inf, deathAge - age))
}

event_death_without_CVD <- function(traj, inputs)
{
  cat("event_Death() ", now(env), "\n")
  
  traj %>% 
    set_attribute("aAgeDeathNonCVD", function() get_attribute(env,"aDeathWithoutCVD") + get_attribute(env,"aAgeInitial")) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death without CVD") %>%
      mark("death_without_CVD") %>%
      terminate(inputs)
  )
}

years_till_get_CVD <- function(inputs)
{
  gender = ifelse(get_attribute(env, "aGender") == 1, "M", "F")
  race_pcr = race_convert(get_attribute(env, "aRace")) 
  
  prob = pcr(gender,
             get_attribute(env,"aAgeInitial"),
             race_pcr,
             get_attribute(env, "aTotChol"),
             get_attribute(env, "aHdlChol"),
             get_attribute(env, "aSystolicBp"),
             (get_attribute(env, "aBpTreatment") == 1),
             (get_attribute(env, "aSmoker") %in% c(1,2)),
             (get_attribute(env, "aDiabetic") == 1),
             get_attribute(env, "aPrsZ")
             )/100
  
  # 10-year probability to 1-year rate
  years = rexp(1,ProbToRate(prob, 10))
  
  return(years)
}

event_get_CVD = function(traj, inputs) {
  cat("event_get_CVD() ", now(env), "\n")

  traj %>%
    set_attribute("aAgeCVD", function() get_attribute(env,"aAgeInitial") + get_attribute(env, "aGetCVD")) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Get CVD") %>%
      mark("get_CVD"))
}

years_till_death_of_ASCVD = function(inputs) {
  
  ageCVD    <- get_attribute(env, 'aGetCVD') + get_attribute(env, "aAgeInitial")
  gender    <- get_attribute(env, 'aGender')
  
  ## getCVD maybe sth really big.
  
  age_max = ifelse(gender == 1, length(gompertz.afterASCVD.parameters$male.shape), length(gompertz.afterASCVD.parameters$female.shape))
  
  if (ageCVD >= age_max - 1) {
    # not able to use ageAtDeath(ageCVD+1, gender)
    years = get_attribute(env, 'aGetCVD')
  } else {
    ## ageCVD under the age limitation -- able to use gomepertz
    # whether ASCVD died in first year happened
    happen = 0
    if (gender == 1) {
      ## men
      if (ageCVD < 65) {
        happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_young, inputs$mortalityFirstYearASCVD_male_young))
        } else {
          happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_old, inputs$mortalityFirstYearASCVD_male_old))
          }
      } else {
        ## women
        if (ageCVD < 65) {
          happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_young, inputs$mortalityFirstYearASCVD_female_young))
          } else {
            happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_old, inputs$mortalityFirstYearASCVD_female_old))
          }
      }
    years = ifelse(happen == 1, get_attribute(env, "aGetCVD") + 1, Inf)
  }
  
  return(years)
}

event_death_of_ASCVD = function(traj, inputs) {
  traj %>%
    set_attribute("aAgeDeathCVD", function() get_attribute(env, "aAgeInitial") + now(env)) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death of first-year ASCVD") %>%
      mark("death_of_ASCVD") %>%
      terminate(inputs)
    )
  }

years_till_death_after_ASCVD <- function(inputs) {
  # whether die of ASCVD(in <= 1 yr)
  if (is.infinite(get_attribute(env, 'aDeathOfASCVD'))) {
  
    ageCVD    <- get_attribute(env, 'aGetCVD') + get_attribute(env, "aAgeInitial")
    gender    <- get_attribute(env, 'aGender')

    deathAge = ageAtDeath_afterASCVD(ageCVD + 1, gender)
  
    # GetCVD + ageInitial maybe a impossible number
    years = deathAge - get_attribute(env, "aAgeInitial")
  
  } else {
    years = Inf
  }
  
  return(years)
}

event_death_after_ASCVD = function(traj, inputs) {
  cat("event_death_with_CVD", now(env), "\n")
  
  traj %>% 
    set_attribute("aAgeDeathCVD", function() get_attribute(env, "aAgeInitial")+now(env)) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death after ASCVD event") %>%
      mark("death_after_ASCVD") %>%
      terminate(inputs)
    )
}

event_registry <- list(
        list(name          = "get CVD",
             attr          = "aGetCVD",
             time_to_event = years_till_get_CVD,
             func          = event_get_CVD,
             reactive      = FALSE),
        list(name          = "Death without CVD",
             attr          = "aDeathWithoutCVD",
             time_to_event = years_till_death_without_CVD,
             func          = event_death_without_CVD,
             reactive      = FALSE),
        list(name          = "Death of ASCVD",
             attr          = "aDeathOfASCVD",
             time_to_event = years_till_death_of_ASCVD,
             func          = event_death_of_ASCVD,
             reactive      = FALSE),
        list(name          = "Death after ASCVD",
             attr          = "aDeathAfterASCVD",
             time_to_event = years_till_death_after_ASCVD,
             func          = event_death_after_ASCVD,
             reactive      = FALSE),
        list(name          = "Halt at time horizon",
             attr          = "aTerminate",
             time_to_event = function(inputs) inputs$vHorizon,
             func          = terminate,
             reactive      = FALSE)
)
## aAgeCVD is not immediately(only get after this happen), but aGetCVD happens at 0.

## in this way there'll be 2 aGetCVD
## because if getCVD is inside the horizon, there'll be repeats. for the first version, they just died/terminated. 
```

Two versions of event_registry -- 

```{r run}
source("DES.R")

env = simmer("CVD")

exec.simulation <- function(inputs)
{
        set.seed(11451428)
        env  <<- simmer("CVD")
        traj <- simulation(env, inputs)
        env %>% create_counters(counters)
        
        env %>%
                add_generator("patient", traj, at(rep(0, inputs$vN)), mon=2) %>%
                run(inputs$vHorizon+1e-6) %>% # Simulate just past horizon
                wrap()
        
        get_mon_arrivals(env, per_resource = T)
}

results <- NULL
attributes <- NULL

run <- exec.simulation(inputs)
at <- arrange(get_mon_attributes(env),name,key,time) #obtain attributes data
if(is.null(results)) { results <- run } else  {results <- rbind(results, run)}
                if(is.null(attributes)) { attributes <- at } else  {attributes <- rbind(attributes, at)}
                rm(run)
                rm(at)
```

method2: can see the expected living years both with and without CVD
method1: only if CVD happens before DeathWithoutCVD

What if DeathWithCVD bigger than DeathWithoutCVD, but the person get CVD before death


```{r results}
attributes
  # as the ageCVD might be over 79, ageCVDagain is not correct. 
  # aGetCVD and aAgeCVD only draw the first time. 

# repeated get CVD, only draw the first time
repeatNames = attributes %>% 
  filter(time != 0 & key == "aGetCVD") %>%
  select(name) %>% unique(.) 

molten = repeatNames %>% left_join(attributes, by = "name") %>%
  filter(key == "aGetCVD" | key == "aAgeCVD") %>%
  group_by(name, key) %>%
  summarize(time = min(time)) %>% 
  left_join(attributes %>% filter(key == "aAgeCVD" | key == "aGetCVD") %>% select(-replication), by = c("name", "key", "time")) %>% 
  rbind(attributes %>% filter(name %in% repeatNames$name & !(key %in% c("aAgeCVD", "aGetCVD"))) %>% select(-replication)) %>%
  rbind(attributes %>% filter(!(name %in% repeatNames$name)) %>% select(-replication)) %>%
  select(-time)

casted = dcast(molten, name ~ key)
casted
```
What if DeathWithCVD bigger than DeathWithoutCVD, but the person get CVD before death
-- it is like that a person get into another distribution(or you can say fate) after getting CVD. 
But for the consistency of 1.9, just use the newest DeathWithCVD. 

die with/die of

```{r question}
# casted %>% filter(aGetCVD < aDeathWithoutCVD & aAgeDeathCVD > aAgeInitial + aDeathWithoutCVD)
```
```{r patient-group}
## 1. never get CVD, never die before horizon. 
casted %>% filter(is.na(aAgeCVD) & is.na(aAgeDeathNonCVD))

## 2. never get CVD, die of background mortality before horizon
casted %>% filter(is.na(aAgeCVD) & !is.na(aAgeDeathNonCVD))

## 3. get CVD, survive till horizon
casted %>% filter(!is.na(aAgeCVD) & is.na(aAgeDeathCVD))

## 4. get CVD, die of first-year ASCVD
casted %>% filter(!is.na(aAgeCVD) & !is.na(aAgeDeathCVD) & aDeathOfASCVD == 1 + aGetCVD)

## 5. get CVD, die after first-year ASCVD event
casted %>% filter(!is.na(aAgeCVD) & !is.na(aAgeDeathCVD) & is.finite(aDeathAfterASCVD))

## exception1: get CVD, die of initial background mortality
casted %>% filter(!is.na(aAgeCVD) & !is.na(aAgeDeathNonCVD))

```

```{r results}
results %>% filter(resource == "time_in_model") %>% mutate(time_in_model = activity_time) %>% select(name, time_in_model)

repeatNames2 = results %>% filter(resource != "time_in_model") %>% count(name, resource) %>% filter(n > 1) %>% pull(name)

attributes %>% filter(name %in% repeatNames2) %>% filter(time > 0)
# there will no repeated death with CVD because the repeat will only happen after the event happen, but for death the event will terminate all. 

trace = results %>% filter(name %in% repeatNames2 & resource == "get_CVD") %>% group_by(name, resource) %>% summarize(start_time = min(start_time)) %>%
  rbind(results %>% filter(!(name %in% repeatNames2) | resource != "get_CVD") %>%
          filter(resource != "time_in_model") %>%
          select(name, resource, start_time)) %>%
  spread(key = resource, value= start_time) %>%
  full_join(results %>% filter(resource == "time_in_model") %>% mutate(time_in_model = activity_time) %>% select(name, time_in_model), by = "name") %>%
  left_join(casted %>% select(name, initial_age = aAgeInitial), by = "name")

# trace %>% filter(get_CVD == death_of_ASCVD)

trace
```

```{r cost-utility}

## 1. never get CVD, never die before horizon. 
## cost: inputs$cHealthcareNonCVD * trace$time_in_model
## payoff: inputs$uHealthy * trace$time_in_model

## 2. never get CVD, die of background mortality before horizon
## cost: inputs$cHealthcareNonCVD * trace$time_in_model
## payoff: inputs$uHealthy * trace$time_in_model

## 3. get CVD, survive till horizon
## cost: inputs$cHealthcareNonCVD * trace$time_in_model + inputs$cNonFatalASCVD + inputs$cAnnualFU_afterASCVD * (trace$time_in_model - trace$get_CVD - 1)
## payoff: inputs$uHealthy * trace$get_CVD + inputs$uAfterASCVD*(trace$time_in_model - trace$get_CVD)

## 4. get CVD, die of first-year ASCVD
## cost: inputs$cHealthcareNonCVD * trace$time_in_model + inputs$cFatalASCVD
## payoff: inputs$uHealthy * trace$get_CVD + inputs$uAfterASCVD*(trace$time_in_model - trace$get_CVD)

## 5. get CVD, die after first-year ASCVD event
## cost: inputs$cHealthcareNonCVD * trace$time_in_model + inputs$cNonFatalASCVD + inputs$cAnnualFU_afterASCVD * (trace$time_in_model - trace$get_CVD - 1)
## payoff: inputs$uHealthy * trace$get_CVD + inputs$uAfterASCVD*(trace$time_in_model - trace$get_CVD)

basic_CVD_costs = function(inputs, death_after_ASCVD, death_of_ASCVD, death_without_CVD, get_CVD, time_in_model, initial_age) {
  cost = 0
  ##cHealthcareNonCVD
  if (initial_age < 45) {
    cost = cost + inputs$cHealthcareNonCVD_18_44 * (min(45, time_in_model + initial_age) - initial_age) + inputs$cHealthcareNonCVD_45_64 * (max(time_in_model + initial_age, 45) - 45)
  } else if (initial_age >= 45 & initial_age < 55) {
    cost = cost + inputs$cHealthcareNonCVD_45_64 * time_in_model
  } else if (initial_age >= 55 & initial_age < 65) {
    cost = cost + inputs$cHealthcareNonCVD_45_64 * (min(65, time_in_model + initial_age) - initial_age) + inputs$cHealthcareNonCVD_65 * (max(time_in_model + initial_age, 65) - 65)
  } else {
    cost = cost + inputs$cHealthcareNonCVD_65 * time_in_model
  }
  
  # FatalASCVD
  if(!is.na(get_CVD) & !is.na(death_of_ASCVD)) {
    cost = cost + inputs$cFatalASCVD
  }
  
  # Non-Fatal ASCVD
  if(!is.na(get_CVD) & is.na(death_of_ASCVD)) {
    cost = cost + inputs$cNonFatalASCVD + inputs$cAnnualFU_afterASCVD * (time_in_model - get_CVD - 1)
  }
  return(cost)
}

trace %>% mutate(cost = basic_CVD_costs(inputs, death_after_ASCVD, death_of_ASCVD, death_without_CVD, get_CVD, time_in_model, initial_age))
```


```{r check}
pcr(ifelse(casted[2,"aGender"] == 1, "M", "F"), 
    casted[2, "aAgeInitial"],
    race_convert(casted[2, "aRace"]),
    casted[2, "aTotChol"],
    casted[2, "aHdlChol"],
    casted[2, "aSystolicBp"],
    casted[2, "aBpTreatment"] == 1,
    casted[2, "aSmoker"] %in% c(1,2),
    casted[2, "aDiabetic"] == 1,
    casted[2, "aPrsZ"]
    )

attributes %>% filter(key == "aAgeDeath") %>% inner_join(attributes %>% filter(key == "aAgeDeathCVD"), by = "name")
attributes %>% filter(key == "aAgeCVD") %>% inner_join(attributes %>% filter(key == "aAgeDeathCVD"), by = "name")
results %>% filter(resource == "time_in_model")

## all the events should count from time 0 
casted %>% filter(aDeathAfterASCVD + aAgeInitial == aAgeDeathCVD)

cHealthcareNonCVD = function(inputs, time_in_model, initial_age){
  cost = 0
  if (initial_age < 45) {
    cost = cost + inputs$cHealthcareNonCVD_18_44 * (min(45, time_in_model + initial_age) - initial_age) + inputs$cHealthcareNonCVD_45_64 * (max(time_in_model + initial_age, 45) - 45)
  } else if (initial_age >= 45 & initial_age < 55) {
    cost = cost + inputs$cHealthcareNonCVD_45_64 * time_in_model
  } else if (initial_age >= 55 & initial_age < 65) {
    cost = cost + inputs$cHealthcareNonCVD_45_64 * (min(65, time_in_model + initial_age) - initial_age) + inputs$cHealthcareNonCVD_65 * (max(time_in_model + initial_age, 65) - 65)
  } else {
    cost = cost + inputs$cHealthcareNonCVD_65 * time_in_model
  }
  return(cost)
}

trace %>%
  mutate(cHealthcareNonCVD = cHealthcareNonCVD(inputs, time_in_model, initial_age))

```