---
title: "Common Random Number in DES simulation -- based on CVD model"
author: "Astrid Yu"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
---

# Preparation

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
load.lib<-c("reshape2","tidyverse","ggrepel","dampack","kableExtra","flexsurv","readxl","haven","janitor","here","parallel")
install.lib<-load.lib[!load.lib %in% installed.packages()]
for(lib in install.lib) install.packages(lib,dependencies=TRUE)
sapply(load.lib,require,character=TRUE)
library(simmer)
select <- dplyr::select

options("scipen"=1000, "digits"=4)
```

## Inputs

Spahillari, et al. (2020)

```{r inputs}
params = list(
  ## basic/general from Stroke paper, 2023
  strategy = 0, # or 1
  vN = 1000,
  vHorizon = 10,
  discount_rate = 0.03,
  wtp = 100000,
  CPI2017 = 245.12,
  CPI2019 = 255.657,
  CPI2023 = 302.408,
  
  ## events from paper 2017
  rrStatinsASCVD = 0.79,
  rrStatinsASCVD_low = 0.77,
  rrStatinsASCVD_upp = 0.81,
  rrStatinsASCVD_dist = "lnorm",
  
  hrAfterASCVD = 1.9,
  hrAfterASCVD_low = 1.60,
  hrAfterASCVD_upp = 2.40, 
  hrAfterASCVD_dist = "lnorm", # log-normal
  
  mortalityFirstYearASCVD_female_young = 0.09,
  mortalityFirstYearASCVD_female_young_dist = "beta",
  
  mortalityFirstYearASCVD_female_old = 0.3,
  mortalityFirstYearASCVD_female_old_dist = "beta",
  
  mortalityFirstYearASCVD_male_young = 0.14,
  mortalityFirstYearASCVD_male_young_dist = "beta",
  
  mortalityFirstYearASCVD_male_old = 0.25,
  mortalityFirstYearASCVD_male_old_dist = "beta",
  
  pMildStatinAdverse = 0.047,
  pMildStatinAdverse_low = 0.047*0.85, 
  pMildStatinAdverse_upp = 0.047*1.15,
  pMildStatinAdverse_dist = "beta",
  
  pMajorStatinAdverse = 0.006/100,
  pMajorStatinAdverse_low = 0.006/100*0.85,
  pMajorStatinAdverse_upp = 0.006/100*1.15,
  pMajorStatinAdverse_dist = "beta",
  
  pMajorStatinAdverseBeingFatal = 0.09/100,
  pMajorStatinAdverseBeingFatal_low = 0.09/100*0.85,
  pMajorStatinAdverseBeingFatal_upp = 0.09/100*1.15,
  pMajorStatinAdverseBeingFatal_dist = "beta",

  ## utility
  uHealthy = 1,
  uAfterASCVD = 0.773,
  uAfterASCVD_low = 0.773*0.85,
  uAfterASCVD_upp = 0.773*1.15,
  uAfterASCVD_dist = "beta",
  
  uHealthyStatin = 0.996,
  uHealthyStatin_low = 0.991,
  uHealthyStatin_upp = 1.000,
  uHealthyStatin_dist = "beta",
  
  uPenaltyMildStatinAdverse = -0.0055,
  uPenaltyMajorStatinAdverse = -0.0383,
  
  uDeath = 0,
  
  ## cost(2017)
  cAnnualFU_afterASCVD = 3917,
  cAnnualFU_afterASCVD_low = 2611,
  cAnnualFU_afterASCVD_upp = 6528,
  cAnnualFU_afterASCVD_dist = "gamma",
  
  cNonFatalASCVD = 49348,
  cNonFatalASCVD_low = 49348*0.85,
  cNonFatalASCVD_upp = 49348*1.15,
  cNonFatalASCVD_dist = "gamma",
  
  cFatalASCVD = 16760,
  cFatalASCVD_low = 16760*0.85,
  cFatalASCVD_upp = 16760*1.15,
  cFatalASCVD_dist = "gamma",
  
  cAnnualStatin = 84,
  
  cMildStatinAdverse = 215,
  cMildStatinAdverse_low = 196,
  cMildStatinAdverse_upp = 326,
  cMildStatinAdverse_dist = "gamma",
  
  cMajorStatinAdverse = 8486,
  cMajorStatinAdverse_low = 6920, 
  cMajorStatinAdverse_upp = 10314,
  cMajorStatinAdverse_dist = "gamma",
  
  annual_discount_rate = 0.03,
  
  ### source：Agency for Healthcare Research and Quality. Mean expenditure per person by age groups, United States, 1996 to 2020. Medical Expenditure Panel Survey.(2020)
  cHealthcareNonCVD_18_44 = 3901,
  cHealthcareNonCVD_45_64 = 8693,
  cHealthcareNonCVD_65 = 12441

)
## inflation
params$rInflation2017 = params$CPI2023 / params$CPI2017
params$rInflation2019 = params$CPI2023 / params$CPI2019

## leave some spaces for PSA
```

```{r func}
# simply change from prob to rate than to event
ProbToRate = function(prob, t){
  -log(1-prob)/t
}

RateToProb = function(rate, t){
  1 - exp(-rate*t)
}
# # what this t should be -- 10 years?
# # time unit: year
# rate = ProbToRate(pcr()/100)
# days = rexp(1,rate)

params$cont_discount_rate <- -log(1- params$annual_discount_rate) # Yearly Time Scale
discounted <- function(undiscounted, start_year, end_year, rate = params$cont_discount_rate)
{
  undiscounted/ (exp(-rate*(end_year-start_year)))
}

discount <- function(value, A, ar=params$annual_discount_rate) value / (1+ar)^A
```

## PCE

We use the pooled cohort equation to calculate the risk of CVD from risk factors.

<https://github.com/spgarbet/desdt/tree/main/framingham>

## Sample population

We use NHANES study for the simulated population (limited by the PCE limits).

Use the 2017-2018 version: <https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2017>

The sample is vascular-history-free:\
MCQ160d - Ever told you had angina/angina pectoris\
MCQ160e - Ever told you had heart attack\
MCQ160f - Ever told you had a stroke

```{r nhanes-clearance, eval=FALSE}
# source("building_nhanes_sample.R")
load("basepop.RData")
```

## Gompertz Model

We use Gompertz model based on US cause-deleted life table to predict the death age.

```{r source}
source("nonCVD_life_table.R")
load("life table/MarkovCVD_gompertz.RData")
```

# PRS

1.  add this Z score PRS rnorm(1,0,1) [-5,5]\
2.  use 1.73 Hazard ratio from Nature Medicine Paper

10-year ASCVD prob adding PRS risk:\
1 - exp(log(1 - pce_orig) \* 1.73\^PRS)

## Landscape

```{r}
# 2018 guidelines high risk -- must use statins
subset(nhanes_sample, LBDLDL >= 190 | DIQ010 == 1 | pce_orig >= 0.2)

# low risk -- needn't need use statins
subset(nhanes_sample, LBDLDL < 190 & DIQ010 != 1 & pce_orig < 0.05)

# borderline risk 0.05-0.075
hist(subset(nhanes_sample,LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.05 & pce_orig < 0.075)$pce_prs)
mean(subset(nhanes_sample,LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.05 & pce_orig < 0.075)$pce_prs > 0.075)
mean(subset(nhanes_sample,LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.05 & pce_orig < 0.075)$pce_prs < 0.05)
mean(subset(nhanes_sample,LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.05 & pce_orig < 0.075)$pce_prs > 0.2)

# Intermediate risk
hist(subset(nhanes_sample,LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.075 & pce_orig < 0.2)$pce_prs)
mean(subset(nhanes_sample,LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.075 & pce_orig < 0.2)$pce_prs > 0.2)
mean(subset(nhanes_sample,LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.075 & pce_orig < 0.2)$pce_prs < 0.05)
```

## Detailed camparison plots

```{r}
# Total
nhanes_sample %>%
  pivot_longer(cols = starts_with("pce_"), names_to = "PCE_type", names_prefix = "pce_", values_to = "prob_10yr") %>%
  ggplot(aes(x=prob_10yr, color = PCE_type)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.02) +
  geom_density() +
  facet_grid(PCE_type ~ .) + 
  labs(title = "Distribution of 10-yr ASCVD Risk (w/o + w/ PRS)") + 
  theme_bw()

## PRS won't change the decision
# High Risk, must use statins 
nhanes_sample %>%
  filter(LBDLDL >= 190 | DIQ010 == 1 | pce_orig >= 0.2) %>%
  pivot_longer(cols = starts_with("pce_"), names_to = "PCE_type", names_prefix = "pce_", values_to = "prob_10yr") %>%
  ggplot(aes(x=prob_10yr, color = PCE_type)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.02) +
  geom_density() +
  facet_grid(PCE_type ~ .) + 
  labs(title = "Distribution of High 10-yr ASCVD Risk (w/o + w/ PRS)") + 
  theme_bw()

# Low Risk, < 5%
nhanes_sample %>%
  filter(LBDLDL < 190 & DIQ010 != 1 & pce_orig < 0.05) %>%
  pivot_longer(cols = starts_with("pce_"), names_to = "PCE_type", names_prefix = "pce_", values_to = "prob_10yr") %>%
  ggplot(aes(x=prob_10yr, color = PCE_type)) +
  geom_histogram(aes(y = ..density..), bins = 30) +
  geom_density() +
  facet_grid(PCE_type ~ .) + 
  labs(title = "Distribution of Low 10-yr ASCVD Risk (w/o + w/ PRS)") + 
  theme_bw()

# PRS may change the decision
## Borderline Risk, 5% - 7.5%
nhanes_sample %>%
  filter(LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.05 & pce_orig < 0.075) %>%
  pivot_longer(cols = starts_with("pce_"), names_to = "PCE_type", names_prefix = "pce_", values_to = "prob_10yr") %>%
  ggplot(aes(x=prob_10yr, color = PCE_type)) +
  geom_histogram(aes(y = ..density..), bins = 30) +
  geom_density() +
  facet_grid(PCE_type ~ .) + 
  labs(title = "Distribution of Borderline 10-yr ASCVD Risk (w/o + w/ PRS)") + 
  theme_bw()

## Intermediate Risk
nhanes_sample %>%
  filter(LBDLDL < 190 & DIQ010 != 1 & pce_orig >= 0.075 & pce_orig < 0.2) %>%
  pivot_longer(cols = starts_with("pce_"), names_to = "PCE_type", names_prefix = "pce_", values_to = "prob_10yr") %>%
  ggplot(aes(x=prob_10yr, color = PCE_type)) +
  geom_histogram(aes(y = ..density..), bins = 30) +
  geom_density() +
  facet_grid(PCE_type ~ .) + 
  labs(title = "Distribution of Intermediate 10-yr ASCVD Risk (w/o + w/ PRS)") + 
  theme_bw()

```

# Intervention

strategy 0. use statins as 2018 ACC/AHA

Individuals(age 40 - 80):

-   with diabetes, or

-   low-density lipoprotein cholesterol level \>= 190 mg/dl, or

-   estimated 10-year ASCVD risk greater than or equal to 7.5%.

Statins will be taken until the ASCVD happen.

Adverse events:

Adverse events will happen immediately after using statins. The major adverse event will overlap the mild adverse event. The death caused by the major adverse event should happen immediately after the event.

Strategy 1. use statins as 2018 ACC/AHA, but use the PRS-revised PCE instead.
