---
title: "Basic Life-Death model in DES simulation"
author: "Astrid Yu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
library(simmer)
library(tidyverse)
library(survival)
library(flexsurv)
library(janitor)
library(readxl)
library(haven)
library(reshape2)
```

# gompertz

qx: Probability of dying between ages x and x + 1
lx: Number  surviving to  age x
dx: Number dying between  ages x and x + 1
Lx: Person-years  lived between  ages x and x + 1
Tx: Total number of person-years lived above age x
ex: Expectation of  life at age x

```{r gompertz}
US_life_female= read.table("./life table/fltper_1x1.txt", header = TRUE, skip = 2)
US_life_male= read.table("./life table/mltper_1x1.txt", header = TRUE, skip = 2)

######## female ########

US_life_female_2010 = US_life_female %>%
  filter(Year == "2010")

# change life table into survival data
dx_female_2010 = data.frame(Age = rep(NA, 100000), Death = rep(1, 100000))

start = 1
for (age in 0:(nrow(US_life_female_2010)-1)) {
  n_row = US_life_female_2010[age+1,"dx"]
  dx_female_2010[seq(start,(n_row + start - 1)),"Age"] = age
  start = start + n_row
}

parameters_f = data.frame(
  Age = US_life_female_2010$Age,
  shape=rep(NA,nrow(US_life_female_2010)),
  rate=rep(NA,nrow(US_life_female_2010)))

for (age in 0:nrow(US_life_female_2010)-1) {
  surv.data <- with(dx_female_2010[dx_female_2010$Age >= age,], Surv(Age, Death, origin=age))
  surv.model <- flexsurvreg(surv.data ~ 1, dist="gompertz")
  
  parameters_f$shape[age+1] <- surv.model$coefficients[1]
  parameters_f$rate[age+1] <- exp(surv.model$coefficients[2])
}

######## male ########

US_life_male_2010 = US_life_male %>%
  filter(Year == "2010")

# change life table into survival data
dx_male_2010 = data.frame(Age = rep(NA, 100000), Death = rep(1, 100000))

start = 1
for (age in 0:(nrow(US_life_male_2010)-1)) {
  n_row = US_life_male_2010[age+1,"dx"]
  dx_male_2010[seq(start,(n_row + start - 1)),"Age"] = age
  start = start + n_row
}

parameters_m = data.frame(
  Age = US_life_male_2010$Age,
  shape=rep(NA,nrow(US_life_male_2010)),
  rate=rep(NA,nrow(US_life_male_2010)))

for (age in 0:nrow(US_life_male_2010)-1) {
  surv.data <- with(dx_male_2010[dx_male_2010$Age >= age,], Surv(Age, Death, origin=age))
  surv.model <- flexsurvreg(surv.data ~ 1, dist="gompertz")
  
  parameters_m$shape[age+1] <- surv.model$coefficients[1]
  parameters_m$rate[age+1] <- exp(surv.model$coefficients[2])
}

gompertz.parameters = list(Age = 0:(nrow(parameters_f)-2), male.shape = parameters_m[1:(nrow(parameters_m)-1),"shape"], male.rate = parameters_m[1:(nrow(parameters_m)-1),"rate"], female.shape = parameters_f[1:(nrow(parameters_f)-1),"shape"], female.rate = parameters_f[1:(nrow(parameters_f)-1),"rate"])

ageAtDeath <- function(currentAge, gender)
{
  currentAge_f <- floor(currentAge) #negative time to secular death issue
  
  # Male 1 FEMALE 2
  if(gender == 2)
  {
    shape <- gompertz.parameters$female.shape[currentAge_f+1]
    rate  <- gompertz.parameters$female.rate[currentAge_f+1]
  }
  else
  {
    shape <- gompertz.parameters$male.shape[currentAge_f+1]
    rate  <- gompertz.parameters$male.rate[currentAge_f+1]
  }
  
  min(currentAge + rgompertz(1, shape, rate), 110)
}
```

```{r hazard-gome}
life_female_afterASCVD = data.frame(Year = rep(2010,111), 
                                    Age = 0:110, 
                                    qx = qx = sapply(sapply(US_life_female_2010$qx, ProbToRate, t = 1) * inputs$mortalityAfterASCVD, RateToProb, t = 1),
                                    lx = rep(0,111), 
                                    dx = rep(NA,111))
life_female_afterASCVD$lx[1] = 100000
for (i in 1:nrow(life_female_afterASCVD)) {
  if (life_female_afterASCVD[i, "lx"] > 0) {
    life_female_afterASCVD[i, "dx"] = round(life_female_afterASCVD[i, "qx"] * life_female_afterASCVD[i, "lx"])
    life_female_afterASCVD[i+1, "lx"] = life_female_afterASCVD[i, "lx"] - life_female_afterASCVD[i, "dx"]
  }
}
life_female_afterASCVD  = life_female_afterASCVD %>% filter(!is.na(dx))

dx_female_afterASCVD = data.frame(Age = rep(NA, 100000), Death = rep(1, 100000))

start = 1
for (age in 0:(nrow(life_female_afterASCVD)-1)) {
  n_row = life_female_afterASCVD[age+1,"dx"]
  dx_female_afterASCVD[seq(start,(n_row + start - 1)),"Age"] = age
  start = start + n_row
}

parameters_f = data.frame(
  Age = life_female_afterASCVD$Age,
  shape=rep(NA,nrow(life_female_afterASCVD)),
  rate=rep(NA,nrow(life_female_afterASCVD)))

for (age in 0:nrow(life_female_afterASCVD)-1) {
  surv.data <- with(dx_female_afterASCVD[dx_female_afterASCVD$Age >= age,], Surv(Age, Death, origin=age))
  surv.model <- flexsurvreg(surv.data ~ 1, dist="gompertz")
  
  parameters_f$shape[age+1] <- surv.model$coefficients[1]
  parameters_f$rate[age+1] <- exp(surv.model$coefficients[2])
}

#### male ####


life_male_afterASCVD = data.frame(Year = rep(2010,111), 
                                  Age = 0:110, 
                                  qx = sapply(sapply(US_life_male_2010$qx, ProbToRate, t = 1) * inputs$mortalityAfterASCVD, RateToProb, t = 1),
                                  lx = rep(0,111), dx = rep(NA,111))
life_male_afterASCVD$lx[1] = 100000
for (i in 1:nrow(life_male_afterASCVD)) {
  if (life_male_afterASCVD[i, "lx"] > 0) {
    life_male_afterASCVD[i, "dx"] = round(life_male_afterASCVD[i, "qx"] * life_male_afterASCVD[i, "lx"])
    life_male_afterASCVD[i+1, "lx"] = life_male_afterASCVD[i, "lx"] - life_male_afterASCVD[i, "dx"]
  }
}
life_male_afterASCVD  = life_male_afterASCVD %>% filter(!is.na(dx))

dx_male_afterASCVD = data.frame(Age = rep(NA, 100000), Death = rep(1, 100000))

start = 1
for (age in 0:(nrow(life_male_afterASCVD)-1)) {
  n_row = life_male_afterASCVD[age+1,"dx"]
  dx_male_afterASCVD[seq(start,(n_row + start - 1)),"Age"] = age
  start = start + n_row
}

parameters_m = data.frame(
  Age = life_male_afterASCVD$Age,
  shape=rep(NA,nrow(life_male_afterASCVD)),
  rate=rep(NA,nrow(life_male_afterASCVD)))

for (age in 0:nrow(life_male_afterASCVD)-1) {
  surv.data <- with(dx_male_afterASCVD[dx_male_afterASCVD$Age >= age,], Surv(Age, Death, origin=age))
  surv.model <- flexsurvreg(surv.data ~ 1, dist="gompertz")
  
  parameters_m$shape[age+1] <- surv.model$coefficients[1]
  parameters_m$rate[age+1] <- exp(surv.model$coefficients[2])
}

gompertz.afterASCVD.parameters = list(Age = 0:(nrow(parameters_f)-2), male.shape = parameters_m[1:(nrow(parameters_m)-1),"shape"], male.rate = parameters_m[1:(nrow(parameters_m)-1),"rate"], female.shape = parameters_f[1:(nrow(parameters_f)-1),"shape"], female.rate = parameters_f[1:(nrow(parameters_f)-1),"rate"])

ageAtDeath_afterASCVD <- function(currentAge, gender)
{
  currentAge_f <- floor(currentAge) #negative time to secular death issue
  
  # Male 1 FEMALE 2
  if(gender == 2)
  {
    shape <- gompertz.afterASCVD.parameters$female.shape[currentAge_f+1]
    rate  <- gompertz.afterASCVD.parameters$female.rate[currentAge_f+1]
  }
  else
  {
    shape <- gompertz.afterASCVD.parameters$male.shape[currentAge_f+1]
    rate  <- gompertz.afterASCVD.parameters$male.rate[currentAge_f+1]
  }
  
  min(currentAge + rgompertz(1, shape, rate), 105)
}
save(gompertz.parameters, gompertz.afterASCVD.parameters, file = "./life table/gompertz.RData")
```

# NHANES dataset preparation

```{r nhanes-read}
# Demographic Variables and Sample Weights
## RIAGENDR - Gender
## 1	Male
## 2	Female	
## .	Missing

## RIDAGEYR - Age in years at screening
## 0 to 79	Range of Values
## 80	80 years of age and over
## .	Missing

## WTMEC2YR - Full sample 2 year MEC exam weight
## 2566.1838545 to 419762.83649	Range of Values
## 0	Not MEC Examined
## .	Missing
nhanes_data_raw = data.frame(read_xpt("./NHANES/DEMO_J.XPT", col_select = c("SEQN","RIDAGEYR","RIAGENDR","WTMEC2YR"))) #WTMEC2YR is the sample weights variable

nhanes_pop = nhanes_data_raw
colnames(nhanes_pop) = c("id","gender","age","wt")
nhanes_pop = na.omit(nhanes_pop)
nhanes_pop = nhanes_pop %>%
  filter(wt != 0)
save(nhanes_pop, file = "./NHANES/nhanes.RData")
```

# DES model

```{r inputs}
inputs = NULL
inputs$vN = 1000
inputs$vHorizon = 120
```

```{r des-func}
source("DES.R")
```

```{r live_death_des}
# there're sequence for the counters
counters = c("death", "time_in_model")

initialize_patient <- function(traj, inputs)
{
        traj %>%
                seize("time_in_model")       %>%
                
                # NHANES version
                set_attribute("aNo", function() sample(1:nrow(nhanes_pop), 1, prob=1/nhanes_pop$wt)) %>%
                set_attribute("aGender",    function() nhanes_pop$gender[get_attribute(env, "aNo")]) %>% 
                set_attribute("aAge",       function() nhanes_pop$age[get_attribute(env, "aNo")]) %>% 
                
                set_attribute("aAgeInitial",function() get_attribute(env, "aAge"))
}

# Given attributes of a patient (trajectory), it returns in days 
# how long till the patient would die a secular death.
#
# NOTE: The variable in must be named attrs
years_till_death <- function(inputs)
{
  age       <- get_attribute(env, 'aAge')
  gender    <- get_attribute(env, 'aGender')
  
  deathAge = ageAtDeath(age, gender)
  
  return(deathAge - age)
}

# Given a trajectory, modify as needed when a secular
# death occurs.
#
# In this case, it marks a counter and terminates 
# the trajectory. A branch is required, even though
# it doesn't branch to force the termination.
terminate <- function(traj, inputs)
{
  traj %>%
  branch(function() 1,
         continue=FALSE,
         trajectory()               %>%
           release("time_in_model") %>%
           timeout(0)
        )
}

event_death <- function(traj, inputs)
{
  cat("event_Death() ", now(env), "\n")

  traj %>% 
    set_attribute("aAge", function() get_attribute(env,"aDeath"), mod = "+") %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death") %>%
      mark("death") %>%
      terminate(inputs)
  )
}

event_registry <- list(
        list(name          = "Death",
             attr          = "aDeath",
             time_to_event = years_till_death,
             func          = event_death,
             reactive      = FALSE)
)


env = simmer("life-death")

exec.simulation <- function(inputs)
{
        set.seed(114514)
        env  <<- simmer("life-death")
        traj <- simulation(env, inputs)
        env %>% create_counters(counters)
        
        env %>%
                add_generator("patient", traj, at(rep(0, inputs$vN)), mon=2) %>%
                run(inputs$vHorizon+1) %>% # Simulate just past horizon
                wrap()
        
        get_mon_arrivals(env, per_resource = T)
}

results <- NULL
attributes <- NULL

run <- exec.simulation(inputs)
at <- arrange(get_mon_attributes(env),name,key,time) #obtain attributes data
if(is.null(results)) { results <- run } else  {results <- rbind(results, run)}
                if(is.null(attributes)) { attributes <- at } else  {attributes <- rbind(attributes, at)}
                rm(run)
                rm(at)
```

# Result

```{r results}
attributes %>%
  filter(time != 0) %>%
  ggplot(aes(value)) +
  geom_histogram() +
  ggtitle("Death Age Distribution in DES")

attributes %>%
  filter(key == "aDeath") %>%
  ggplot(aes(value)) +
  geom_histogram() +
  ggtitle("Years Alive Distribution in DES")

molten = attributes %>%
  filter(time == 0) %>%
  select(name, key, value)
casted = dcast(molten, name ~ key) %>% mutate(aAge = aAgeInitial + aDeath)

casted
```

Small question: will create 2 aAge attribute, but I just want to update the aAge
how can they get_attribute if 2?
I can still set a Age only for the newest event, but it will add more ifelse. 


set some normal men/women at different age groups to get LE.
survival curve

```{r verification}
US_life_female_2010[1:110,] %>%
  mutate(age = as.numeric(Age)) %>%
  ggplot(aes(x = age, y = 100000 - lx)) +
  geom_point()

casted %>%
  filter(aGender == 2) %>%
  ggplot(aes(x = aAge)) +
  geom_histogram()

# Use US life table
surv.data <- dx_male_2010[dx_male_2010$Age >= 40,]

hist(surv.data$Age, freq=FALSE, main="Census Data from 2010", xlab="Male Deaths Starting at 40", breaks=20)
curve(dgompertz(x-40, gompertz.parameters$male.shape[41], gompertz.parameters$male.rate[41]) , add=TRUE, col='red')

surv.data <- dx_male_2010[dx_male_2010$Age >= 90,]

hist(surv.data$Age, freq=FALSE, main="Census Data from 2010", xlab="Male Deaths Starting at 90", breaks=20)
curve(dgompertz(x-90, gompertz.parameters$male.shape[91], gompertz.parameters$male.rate[91]) , add=TRUE, col='red')


# reset initial patient

initialize_patient <- function(traj, inputs)
{
        traj %>%
                seize("time_in_model")       %>%
                
                # NHANES version
                set_attribute("aGender",    function() 1) %>% 
                set_attribute("aAge",       function() 40) %>% 
                
                set_attribute("aAgeInitial",function() get_attribute(env, "aAge"))
}
surv.data <- casted

hist(surv.data$aAge, freq=FALSE, main="DES results, 1000 men starting from 40", xlab="Male Deaths Starting at 40", breaks=20)
curve(dgompertz(x-40, gompertz.parameters$male.shape[41], gompertz.parameters$male.rate[41]) , add=TRUE, col='red')


```

# No longer used

```{r events-no-longer-used}
counters = c("get_CVD", "death_without_CVD", "death_with_CVD", "time_in_model")

terminate <- function(traj, inputs)
{
  traj %>%
  branch(function() 1,
         continue=FALSE,
         trajectory()               %>%
           release("time_in_model") %>%
           timeout(0)
        )
}

years_till_death_without_CVD <- function(inputs)
{
  age       <- get_attribute(env, 'aAge')
  gender    <- get_attribute(env, 'aGender')
  
  deathAge = ageAtDeath(age, gender)
  
  return(deathAge - age)
}

event_death_without_CVD <- function(traj, inputs)
{
  cat("event_Death() ", now(env), "\n")

  traj %>% 
    set_attribute("aAgeDeathNonCVD", function() get_attribute(env,"aDeathWithoutCVD") + get_attribute(env,"aAgeInitial")) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death without CVD") %>%
      mark("death_without_CVD") %>%
      terminate(inputs)
  )
}

years_till_get_CVD <- function(inputs)
{
  gender = ifelse(get_attribute(env, "aGender") == 1, "M", "F")
  race_pcr = race_convert(get_attribute(env, "aRace")) 
  
  prob = pcr(gender,
             get_attribute(env,"aAgeInitial") + now(env),
             race_pcr,
             get_attribute(env, "aTotChol"),
             get_attribute(env, "aHdlChol"),
             get_attribute(env, "aSystolicBp"),
             (get_attribute(env, "aBpTreatment") == 1),
             (get_attribute(env, "aSmoker") %in% c(1,2)),
             (get_attribute(env, "aDiabetic") == 1),
             get_attribute(env, "aPrsZ")
             )/100
  
  # 10-year probability to 1-year rate
  years = rexp(1,ProbToRate(prob, 10))
  
  return(years)
}

event_get_CVD = function(traj, inputs) {
  cat("event_get_CVD() ", now(env), "\n")

  traj %>%
    set_attribute("aAgeCVD", function() get_attribute(env,"aAgeInitial") + get_attribute(env, "aGetCVD")) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Get CVD") %>%
      mark("get_CVD")) %>% 
    timeout(function() years_till_death_with_CVD(inputs)) %>%
    event_death_with_CVD(inputs)
}

# years_till_acute_CVD_death = function(inputs) {
#   
#     
#   
#   return(ifelse(happen == 1, 1, 500))
#   }
# 
# event_death_acute_CVD = function(traj, inputs) {
#   traj %>% 
#     set_attribute("aAgeDeath", function() 1 + get_attribute(env,"aAgeCVD")) %>%
#     branch(
#     function() 1,
#     continue=c(TRUE),
#     trajectory("Death Acute CVD") %>%
#       mark("death Acute CVD") %>%
#       terminate(inputs)
#     )
#   }

# years_till_death_with_CVD <- function(inputs)
# {
#   ageCVD    <- get_attribute(env, 'aAgeCVD')
#   gender    <- get_attribute(env, 'aGender')
#   
#   # whether ASCVD died in first year happened
#   happen = 0
#   if (gender == 1) {
#     ## men
#     if (ageCVD < 65) {
#       happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_young, inputs$mortalityFirstYearASCVD_male_young))
#     } else {
#       happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_old, inputs$mortalityFirstYearASCVD_male_old))
#       }
#     } else {
#     ## women
#       if (ageCVD < 65) {
#         happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_young, inputs$mortalityFirstYearASCVD_female_young))
#       } else {
#         happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_old, inputs$mortalityFirstYearASCVD_female_old))
#       }
#     }
#   
#   deathAge = deathAge = ageAtDeath_afterASCVD(
#     ifelse(ageCVD+1>length(gompertz.afterASCVD.parameters$male.shape)-1, length(gompertz.afterASCVD.parameters$male.shape)-1, ageCVD), 
#   gender)
#   
#   return(ifelse(happen == 1, get_attribute(env, "aGetCVD") + 1, get_attribute(env, "aGetCVD") + deathAge - ageCVD))
# }

years_till_death_with_CVD <- function(inputs)
{
  ageCVD    <- get_attribute(env, 'aGetCVD') + get_attribute(env, "aAgeInitial")
  gender    <- get_attribute(env, 'aGender')
  
  ## getCVD maybe sth really big.
  
  age_max = ifelse(gender == 1, max(parameters_m_hr$Age), max(parameters_f_hr$Age))
  
  if (ageCVD >= age_max - 1) {
    # not able to use ageAtDeath(ageCVD+1, gender)
    years = get_attribute(env, 'aGetCVD')
  } else {
    ## ageCVD under the age limitation -- able to use gomepertz
    # whether ASCVD died in first year happened
    happen = 0
    if (gender == 1) {
      ## men
      if (ageCVD < 65) {
        happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_young, inputs$mortalityFirstYearASCVD_male_young))
        } else {
          happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_male_old, inputs$mortalityFirstYearASCVD_male_old))
          }
      } else {
        ## women
        if (ageCVD < 65) {
          happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_young, inputs$mortalityFirstYearASCVD_female_young))
          } else {
            happen = sample(0:1, 1, prob = c(1-inputs$mortalityFirstYearASCVD_female_old, inputs$mortalityFirstYearASCVD_female_old))
          }
        }

  deathAge = ageAtDeath_afterASCVD(ageCVD + 1, gender)
  
  # GetCVD + ageInitial maybe a impossible number
  years = ifelse(happen == 1, get_attribute(env, "aGetCVD") + 1, deathAge - get_attribute(env, "aAgeInitial"))
  }
  
  return(years)
}

event_death_with_CVD = function(traj, inputs) {
  traj %>% 
    set_attribute("aAgeDeathCVD", function() get_attribute(env, "aAgeInitial")+now(env)) %>%
    branch(
    function() 1,
    continue=c(TRUE),
    trajectory("Death With CVD") %>%
      mark("death_with_CVD") %>%
      terminate(inputs)
    )
}

event_registry <- list(
        list(name          = "get CVD",
             attr          = "aGetCVD",
             time_to_event = years_till_get_CVD,
             func          = event_get_CVD,
             reactive      = FALSE),
        list(name          = "Death without CVD",
             attr          = "aDeathWithoutCVD",
             time_to_event = years_till_death_without_CVD,
             func          = event_death_without_CVD,
             reactive      = FALSE),
        # list(name          = "Death with CVD",
        #      attr          = "aDeathWithCVD",
        #      time_to_event = years_till_death_with_CVD,
        #      func          = event_death_with_CVD,
        #      reactive      = FALSE),
        list(name          = "Halt at time horizon",
             attr          = "aTerminate",
             time_to_event = function(inputs) inputs$vHorizon,
             func          = terminate,
             reactive      = FALSE)
)

```